\textbf{ Interpretable Gaussian Processes for Stellar Variability\\ I: Static Surfaces }

### 2020-12-18 tex/ms.tex
more text, last test

\xxx{An example using the 1d Beta distribution.}

@@

\subsection{With limb darkening}
% http://dx.doi.org/10.1051/0004-6361/200913675
%
\label{sec:ld}
We must adjust our expression for the flux in the presence of limb darkening.
For any polynomial limb darkening law of the form
%
\begin{align}
    \label{eq:ld:I}
    \frac{I(\mu)}{I(\mu = 1)} = 1 - \sum_{n=1}^{n_\mathrm{max}} u_n(1 - \mu)^n
    \quad,
\end{align}
%
where $I$ is the intensity on the stellar surface,
$\mu = z = \sqrt{1 - x^2 - y^2}$ is the radial coordinate on the
projected disk, and $u_n$ is a limb darkening coefficient, the effect of
limb darkening on the stellar map can be expressed exactly as a linear
operation on the spherical harmonic coefficient vector
\citep{Luger2019}.
This includes the popular linear and quadractic limb darkening laws
and generalizes to \emph{any} limb darkening law in the limit
$n_\mathrm{max} \rightarrow \infty$. The linearity of the problem
can be understood by noting that all terms
in Equation~(\ref{eq:ld:I}) are strictly polynomials in $x$, $y$, and $z$,
all of which can be expressed exactly as sums of spherical harmonics
\citep{Luger2019}. When weighting the surface intensity by the limb darkening
profile, the resulting intensity is simply a product of spherical harmonics,
which is itself a linear combination of spherical harmonics.
Thus, given a limb darkening law
of degree $n_\mathrm{max}$ with coefficients $\mathbf{u}$,
we can construct a matrix $\mathbf{L}(\mathbf{u})$
that transforms a spherical harmonic vector $\mathbf{y}$ of degree
$l_\mathrm{max}$ to a limb-darkened spherical harmonic vector $\mathbf{y'}$
of degree $l_\mathrm{max} + n_\mathrm{max}$. As an example, consider
a map of degree $l_\mathrm{max} = 1$ and the linear limb darkening law
($n_\mathrm{max} = 1$) with coefficient vector $\mathbf{u} = ( u_1 )$.
The transformation matrix from $\mathbf{y}$ to $\mathbf{y'}$ is
%
\begin{proof}{}
    \label{eq:ld:L}
    \setstackgap{L}{1.5\baselineskip}
    \mathbf{L}(\mathbf{u}) =
    \frac{1}{1 - \frac{u_1}{3}}
    \begin{pmatrix}
        \quadquad1-u_1\quadquad & 0                       & \frac{u_1}{\sqrt{3}}    & 0                       \\
        0                       & \quadquad1-u_1\quadquad & 0                       & 0                       \\
        \frac{u_1}{\sqrt{3}}    & 0                       & \quadquad1-u_1\quadquad & 0                       \\
        0                       & 0                       & 0                       & \quadquad1-u_1\quadquad \\
        0                       & 0                       & 0                       & 0                       \\
        0                       & \frac{u_1}{\sqrt{5}}    & 0                       & 0                       \\
        0                       & 0                       & \frac{2u_1}{\sqrt{15}}  & 0                       \\
        0                       & 0                       & 0                       & \frac{u_1}{\sqrt{5}}    \\
        0                       & 0                       & 0                       & 0
    \end{pmatrix}
    \quad.
\end{proof}
%
The columns of $\mathbf{L}$ are constructed from the coefficient vectors of
each transformed spherical harmonic, which are in turn computed by
multiplying each spherical harmonic by the spherical harmonic representation
of the particular limb darkening law.

>>>>

\subsection{Setup}
\label{sec:calibration-setup}

\begin{table}[t!]
    \begin{center}
        \begin{longtable}{W{c}{1cm} W{l}{6cm} W{l}{2cm}}
            \label{tab:synthetic}
            \\
            %
            \toprule
            \multicolumn{1}{c}{\textbf{Symbol}}
             &
            \multicolumn{1}{c}{\textbf{Description}}
             &
            \multicolumn{1}{c}{\textbf{Value}}
            \\
            \midrule
            \endhead
            \bottomrule                                 \\
            %
            \caption{%
                Default parameters used to generate synthetic light curves in
                the calibration tests.
            }
            \endfoot
            %
            %
            $N$
             & number of spots
             & $\sim\mathcal{N}(20, 0^2)$
            %
            \\
            %
            $c$
             & spot contrast
             & $\sim\mathcal{N}(0.05, 0^2)$
            %
            \\
            %
            $\phi$
             & spot latitude
             & $\sim\mathcal{N}(30^\circ, {5^\circ}^2)$
            %
            \\
            $\lambda$
             & spot longitude
             & $\sim\mathcal{U}(0^\circ, 360^\circ)$
            %
            \\
            %
            $r$
             & spot radius
             & $\sim\mathcal{N}(15^\circ, {0^\circ}^2)$
            %
            \\
            %
            $I$
             & stellar inclination
             & $\sim\sin$
            %
            \\
            %
            $P$
             & rotational period
             & $1$ day
            %
            \\
            %
            $\mathbf{u}$
             & limb darkening coefficients
             & $\left( 0 \,\,\,\, 0 \right)^\top$
            %
            \\
            %
            $\sigma_f$
             & photometric uncertainty
             & $10^{-3}$
            %
            \\
            %
            $K$
             & number of cadences per light curve
             & $10^3$
            %
            \\
            %
            $\Delta t$
             & time baseline
             & $4$ days
            %
            \\
            %
            $M$
             & number of light curves in ensemble
             & $50$
            %
            \\
            %
        \end{longtable}
    \end{center}
\end{table}

We seek to demonstrate that our model is correctly calibrated by testing it
on synthetic data, which we generate as follows. For each of $M$ synthetic
light curves in a given ensemble,
we create a rectangular
$(150 \times 300)$ latitude-longitude grid of surface intensity values, all
intialized
to zero. We then add $N$ spots to this grid, each of fractional contrast
$c$ and radius $r$ centered at latitude $\phi$ and longitude $\lambda$,
by decreasing the intensity at all points within an angular distance $r$
(measured along the surface of the sphere) by an amount $c$.
In order to compute the corresponding light curve, we expand the surface
in spherical harmonics, although at much higher degree
($l_\mathrm{max}^{(0)} = 30$)
than the degree we will use in the inference step ($l_\mathrm{max} = 15$)
to minimize potential ringing effects or other artefacts in the synthetic
data. For reference, the chosen degree $l_\mathrm{max} = 30$ is large
enough to resolve features on the order of
$\nicefrac{180^\circ}{30} = 6^\circ$ across, but small enough that
the algorithm for computing the light curve is numerically stable. We
compute the light curve at $K$ points equally spaced over a baseline $\Delta t$
using the \starry algorithm
(Appendix \ref{sec:starry}), assuming an inclination $I$, a rotational period $P$,
and limb darkening coefficients $\mathbf{u}$. Finally, we divide the
light curve by the mean and add Gaussian noise with standard deviation $\sigma_f$
to emulate photon noise.%
\footnote{
    In theory, we should do this in the reverse order: i.e., we should add photon
    noise and \emph{then} normalize the light curve to the mean, since that is the order in which
    those steps occur in reality. However, if we did that, we would have to normalize $\sigma_f$
    in our inference step, such that the variance of each of the normalized light curves
    in the ensemble would be different, requiring us to invert a different matrix for
    each light curve when computing the log likelihood (see Equation~\ref{eq:log-like}).
    This would significantly increase the computational cost of our tests. Fortunately, in
    practice, the difference between these two approaches has negligible effect on
    our inference results, so we opt for the faster of the two methods. Note, of course,
    that when applying our GP to real data, we won't have this choice!
}
The default values of each of the parameters mentioned above are given in
Table~\ref{tab:synthetic}. Some, like the number of spots, their contrast,
etc., are drawn from fiducial distributions, while others, like the
photometric uncertainty, the rotational period, and the limb darkening
coefficients,
are fixed across all $M$ light curves in an ensemble. These fixed values
are not realistic, but they greatly speed up the inference step, since they
allow us to invert a single covariance matrix for all light curves when computing
the log likelihood.

@@

\subsection{With limb darkening}
% http://dx.doi.org/10.1051/0004-6361/200913675
%
\label{sec:ld}
We must adjust our expression for the flux in the presence of limb darkening.
For any polynomial limb darkening law of the form
%
\begin{align}
    \label{eq:ld:I}
    \frac{I(\mu)}{I(\mu = 1)} = 1 - \sum_{n=1}^{n_\mathrm{max}} u_n(1 - \mu)^n
    \quad,
\end{align}
%
where $I$ is the intensity on the stellar surface,
$\mu = z = \sqrt{1 - x^2 - y^2}$ is the radial coordinate on the
projected disk, and $u_n$ is a limb darkening coefficient, the effect of
limb darkening on the stellar map can be expressed exactly as a linear
operation on the spherical harmonic coefficient vector
\citep{Luger2019}.
This includes the popular linear and quadractic limb darkening laws
and generalizes to \emph{any} limb darkening law in the limit
$n_\mathrm{max} \rightarrow \infty$. The linearity of the problem
can be understood by noting that all terms
in Equation~(\ref{eq:ld:I}) are strictly polynomials in $x$, $y$, and $z$,
all of which can be expressed exactly as sums of spherical harmonics
\citep{Luger2019}. When weighting the surface intensity by the limb darkening
profile, the resulting intensity is simply a product of spherical harmonics,
which is itself a linear combination of spherical harmonics.
Thus, given a limb darkening law
of degree $n_\mathrm{max}$ with coefficients $\mathbf{u}$,
we can construct a matrix $\mathbf{L}(\mathbf{u})$
that transforms a spherical harmonic vector $\mathbf{y}$ of degree
$l_\mathrm{max}$ to a limb-darkened spherical harmonic vector $\mathbf{y'}$
of degree $l_\mathrm{max} + n_\mathrm{max}$. As an example, consider
a map of degree $l_\mathrm{max} = 1$ and the linear limb darkening law
($n_\mathrm{max} = 1$) with coefficient vector $\mathbf{u} = ( u_1 )$.
The transformation matrix from $\mathbf{y}$ to $\mathbf{y'}$ is
%
\begin{proof}{test_ld}
    \label{eq:ld:L}
    \setstackgap{L}{1.5\baselineskip}
    \mathbf{L}(\mathbf{u}) =
    \frac{1}{1 - \frac{u_1}{3}}
    \begin{pmatrix}
        \quadquad1-u_1\quadquad & 0                       & \frac{u_1}{\sqrt{3}}    & 0                       \\
        0                       & \quadquad1-u_1\quadquad & 0                       & 0                       \\
        \frac{u_1}{\sqrt{3}}    & 0                       & \quadquad1-u_1\quadquad & 0                       \\
        0                       & 0                       & 0                       & \quadquad1-u_1\quadquad \\
        0                       & 0                       & 0                       & 0                       \\
        0                       & \frac{u_1}{\sqrt{5}}    & 0                       & 0                       \\
        0                       & 0                       & \frac{2u_1}{\sqrt{15}}  & 0                       \\
        0                       & 0                       & 0                       & \frac{u_1}{\sqrt{5}}    \\
        0                       & 0                       & 0                       & 0
    \end{pmatrix}
    \quad.
\end{proof}
%
The columns of $\mathbf{L}$ are constructed from the coefficient vectors of
each transformed spherical harmonic, which are in turn computed by
multiplying each spherical harmonic by the spherical harmonic representation
of the particular limb darkening law.
